---
title: "PostGIS"
author: "Mirko Mälicke"
output:
  html_notebook:
    theme: flatly
    toc: yes
  html_document:
    theme: flatly
    toc: yes
  pdf_document:
    toc: yes
---

```{r, echo=FALSE, include=T}
# require the package
if (!("RPostgreSQL" %in% installed.packages())){
  install.packages("RPostgreSQL")
}
if (!("getPass" %in% installed.packages())){
  install.packages("getPass")
}
require(RPostgreSQL)
require(getPass)
require(ggplot2)

# establish the connection
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, host='openhydro.de', port=5432, user=getPass('Provide the user'), 
                 password=getPass('Provide the password'), dbname='dwd')
```

Select the biggest catchment from the einzugsgebiete table.
```{sql connection=con}
select id, name, st_area(st_asewkt(geometry)) / 1000000 as "area [km2]" 
from einzugsgebiete 
order by "area [km2]" desc limit 1
```

Select the biggest catchment from the einzugsgebiete table, whichs' name starts with an 'S'.
```{sql connection=con}
select id, name, st_area(st_asewkt(geometry)) / 1000000 as "area [km2]" 
from einzugsgebiete 
where name like 'S%' 
order by "area [km2]" desc limit 1
```

Select the smallest catchment fron the einzugsgebiete table, which contains a station.
```{sql connection=con}
select id, name_catchment, name_station, st_area(st_asewkt(catchment_area)) /1000000 as "catchment area [km2]"
from 
  (select id, name as name_catchment, geometry as catchment_area from einzugsgebiete) e
  join 
  (select name as name_station, geometry as station_coords from stations) s
on st_within(station_coords, catchment_area)
order by "catchment area [km2]" asc limit 1

```


Select the two stations which are closest / farest away from your own location in a single <span style="color:blue">UNION</span> statement.
```{sql connection=con}
select 'closest' as "closest/farest", id, name, distance / 1000 as "distance [km]" from 
  (select st_distance(st_transform(geometry, 25832), (select st_transform(geometry,25832)
  from stations where id=12)) as distance, id, name 
  from stations 
  where id!=12 and geometry is not null 
  order by distance asc limit 1) a
union
select 'farest' as "closest/farest", id, name, distance / 1000 as "distance [km]" from
  (select st_distance(st_transform(geometry, 25832), (select st_transform(geometry,25832) from stations where id=12)) as distance, id, name from stations 
  where id!=12 and geometry is not null 
  order by distance desc limit 1) a



```

In a new project, you would like to build a new weather station either close to the Herderbau or near the Münster in Freiburg. You decide to chose the 
location which has less other stations in a 1.5 km radius
```{sql connection=con}
select 'Münster' as location, count(id) as amount_of_stations
from stations 
where st_within(st_transform(geometry, 25832), st_buffer(st_transform(st_setsrid(st_point(7.852789, 47.995581),4326), 25832),1500)) 
union
select 'Herderbau' as location, count(id) as amount_of_stations
from stations 
where st_within(st_transform(geometry, 25832), st_buffer(st_transform(st_setsrid(st_point(7.85331, 48.0028),4326), 25832),1500))

```

In class, you joined the vegetation class table with its lookup table and calculated the areas for each feature. As you might have noticed, there are duplicates in the description field. 
This is obviously the reason why a lookup table was used. Aggregate these classes together and calculate the overall area for each cover class.
```{sql connection=con}
select name, description_id, sum(area) / 1000000 as "area [km²]" from
  (select vc.id, vc.description_id, d.name, st_area(vc.geometry) as area from
    (select id, description_id, geometry from vegetation_cover) vc
    join vegetation_cover_description d 
    on vc.description_id=d.id
  ) a

group by name, description_id
```

Redo the last step but query the data with nice looking values like '34.1 km²' for the area and add a ratio field, which should give the share of each class on the total area in a '20.1%' style.
```{sql connection=con}
create temporary table share_lk as 
 select name, description_id, round(cast(area as numeric),1) as area from
  (select name, description_id, sum(area) / 1000000 as area from
    (select vc.id, vc.description_id, d.name, st_area(vc.geometry) as area from
      (select id, description_id, geometry from vegetation_cover) vc
      join vegetation_cover_description d 
      on vc.description_id=d.id) a
  group by name, description_id) b
```  
```{sql connection=con}  
select name, description_id, area || 'km²' as area, round(cast((area/(select sum(area) from share_lk))*100 as numeric),1) || '%' as ratio from share_lk
```


Do the einzugsgebiete and vegetation_cover tables cover the same area?
```{sql connection=con}
 
    select 'catchment areas' as "type", st_area(st_union(geometry))/100000 as "area [km²]" from einzugsgebiete 
    union
    select 'vegetation' as "type", st_area(st_union(geometry))/100000 as "area [km²]" from vegetation_cover
```


Select the 10 stations closest to your HOBO location and query their name and geometry. 
Calculate the distance to your HOBO location in different coordinate systems. Use ETRS89/UTM 32N; DHDN/GK Z3 and WGS84/Pseudomercartor.
Do the distances differ and if yes, why?
```{sql connection=con}
--UTM zone 32N
select name, st_asewkt(st_transform(geometry, 25832)) as geometry, 
    st_distance(st_transform(geometry, 25832), (select st_transform(geometry, 25832) from     stations where id = 12)) / 1000 as "distance [km]" 
from stations where id != 12 
order by "distance [km]" asc limit 10;
```
```{sql connection=con}
--DHDN / 3-degree Gauss-Kruger zone 3
select name, st_asewkt(st_transform(geometry, 31467)) as geometry, 
    st_distance(st_transform(geometry, 31467), (select st_transform(geometry, 31467) from     stations where id = 12)) / 1000 as "distance [km]" 
from stations where id != 12 
order by "distance [km]" asc limit 10;
```
```{sql connection=con}
--WGS84 Web Mercator
select name, st_asewkt(st_transform(geometry, 3857)) as geometry, 
    st_distance(st_transform(geometry, 3857), (select st_transform(geometry, 3857) from     stations where id = 12)) / 1000 as "distance [km]" 
from stations where id != 12 
order by "distance [km]" asc limit 10;
```
<div class="alert alert-info">
Answer: They differ, because the different srid were used describing different coordinate reference system. There is no big difference between the used UTM zone 32N and DHDN / 3-degree Gauss-Kruger zone 3, because they are to describe smaller regions, while the WGS84 Web Mercator is made to use it for large-scale maps e.g. world maps. WGS84 Web Mercator is therefore the standard for web applications like google maps.
</div>



### cleanup

```{r}
dbDisconnect(con)
```

