---
title: "SQL Exercises"
author: "Mirko Mälicke"
output: html_notebook
---

This notebook collects some basic SQL exercises for the sql_introduction.Rmd lecture. You can solve them and create a pull request for your soultion in the master Github branch.

```{r, include=FALSE}
# require the package
if (!("RPostgreSQL" %in% installed.packages())){
  install.packages("RPostgreSQL")
}
require(RPostgreSQL)
if (!("getPass" %in% installed.packages())){
  install.packages("getPass")
}
require(RPostgreSQL)
require(getPass)

# establish the connection
drv <- dbDriver('PostgreSQL')
con <- dbConnect(drv, host='openhydro.de', port=5432, user=getPass('Provide the user'), 
                 password=getPass('Provide the password'), dbname='dwd')
```


### SELECT

Select all stations located lower than 400 meter.
```{sql connection=con}
Select name from stations where elevation >400
```

How many stations start with the letter 'W'?
```{sql connection=con}
select count(*) from stations where name like 'W%'

```

Are there more stations starting with 'M' or 'W'?
```{sql connection=con}
select 'W' as letter, count(*) from stations where name like 'W%'
union
select 'M' as letter, count(*) from stations where name like 'M%'
```

### SELECT & Grouping

Find the 5 starting letters in station names with the most occurences.
```{sql connection=con}

select substr(name, 1,1) as first_letter, count(name) from stations group by first_letter order by count desc limit 5
```

Find the 10 starting letters in station names with the lowest mean elevation.
```{sql connection=con}
select substr(name, 1,1) as first_letter, avg(elevation) as elevation, count(name) from stations group by first_letter order by elevation asc limit 10
```

You also learned about the vegetation_cover table. The *description_id* attribute was highly repetitive. But how does the distribution look like between these labels? Count the occurences per description_id.

```{sql connection=con}
select description_id, count(description_id) from vegetation_cover group by description_id

```

Group the vegetation_cover table by the name in vegetation_cover_description and count the instances per label.

```{sql connection=con}
explain analyze select d.name, count(v.id) from vegetation_cover as v
  join vegetation_cover_description as d
  on d.id=v.description_id
  group by d.name
limit 10

```
A similar approach to the problem but writing it differently

```{sql connection=con}
explain analyze select name, count(vegetation_cover.id) from vegetation_cover 
join vegetation_cover_description
on vegetation_cover.description_id=vegetation_cover_description.id group by name limit 20
```
An alternative way of solving the problem (but no the best one) that gives back the same result:
```{sql connection=con}
explain analyze select name, count(*) from vegetation_cover, vegetation_cover_description 
where vegetation_cover.description_id=vegetation_cover_description.id group by name limit 20

```

The three different codes take a similar time to run, so there is no one better option among them.  
```{sql connection=con}
select name, count(vegetation_cover.id) from vegetation_cover 
join vegetation_cover_description
on vegetation_cover.description_id=vegetation_cover_description.id group by name limit 20
```
### cleanup
```{r, include=FALSE}
dbDisconnect(con)
```

